### 2023-10-20 appendix.Rmd
starting writing manuscript

\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}

# **Supplementary Information**

\setcounter{page}{1}
\centering
\noindent \hspace*{10mm} \small Mapping the punishment strategy space \newline
\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, Nichola Raihani\textsuperscript{1,2} \newline
\raggedright
\noindent \footnotesize \textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand \newline
\noindent \footnotesize \textsuperscript{2} Department of Experimental Psychology, University College London, London, United Kingdom \newline
\normalsize
\newpage

## Supplementary Figures

## Supplementary Tables

## Supplementary References