### 2024-07-24 appendix.Rmd
submit manuscript

\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}

# **Supplementary Material**

\setcounter{page}{1}
\centering
\noindent \hspace*{25mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
\hspace*{30mm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, & Nichola Raihani\textsuperscript{1,2} \newline
\raggedright
\noindent \footnotesize \textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand \newline
\noindent \footnotesize \textsuperscript{2} Department of Experimental Psychology, University College London, London, United Kingdom \newline
\par
\noindent{\bfseries{Contents}}
\contentsline {subsection}{Supplementary Figures S1-S14}{2}{}
\contentsline {subsection}{Supplementary Tables S1-S4}{16}{}
\normalsize
\newpage

## Supplementary Figures

(ref:plotModel1aCaption) *Posterior estimates of the probabilities of following
different punishment strategies from the Bayesian latent state model fitted to
the full dataset without pre-registered exclusions.* The model assumes an 
implementation error rate of 5%. Figure \@ref(fig:plotModel1b) in the main text
shows the same result, but from a model fitted to the reduced dataset with 
pre-registered exclusions. Points represent posterior medians, line ranges 
represent 50% and 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

```{r plotModel1a, fig.width=7, fig.height=5, fig.cap="(ref:plotModel1aCaption)", warning=FALSE}
tar_read(plot1.1)
```

\newpage

(ref:plotAllDems2Caption) *Posterior slopes from models including age, 
socio-economic status, gender, and student status, fitted to the subsetted 
dataset with pre-registered exclusions.* Each row represents a separate model. 
Points represent posterior medians, line ranges represent 95% credible
intervals. AI = advantageous inequity, DI = disadvantageous inequity.

```{r plotAllDems2, fig.width=7.5, fig.height=3, fig.cap="(ref:plotAllDems2Caption)", warning=FALSE}
tar_read(plotAllDems2)
```

\newpage

(ref:plotAllDems1Caption) *Posterior slopes from models including age, 
socio-economic status, gender, and student status, fitted to the full dataset 
without pre-registered exclusions.* Each row represents a separate model. Points 
represent posterior medians, line ranges represent 95% credible intervals. AI = 
advantageous inequity, DI = disadvantageous inequity.

```{r plotAllDems1, fig.width=7.5, fig.height=3, fig.cap="(ref:plotAllDems1Caption)", warning=FALSE}
tar_read(plotAllDems1)
```

\newpage

(ref:plotAllPers1Caption) *Posterior slopes from models including Big-6 
personality dimensions and Social Value Orientation, fitted to the full dataset 
without pre-registered exclusions.* Each row represents a separate model. Figure 
\@ref(fig:plotAllPers2) in the main text shows the same results, but from models
fitted to the reduced dataset with pre-registered exclusions. Points represent
posterior medians, line ranges represent 95% credible intervals. AI = 
advantageous inequity, DI = disadvantageous inequity.

```{r plotAllPers1, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPers1Caption)", warning=FALSE}
tar_read(plotAllPers1)
```

\newpage

(ref:plotAllPolRel1Caption) *Posterior slopes from models including political 
ideology, views about social inequality, and religiosity, fitted to the full 
dataset without pre-registered exclusions.* Each row represents a separate model 
aside from Social Dominance Orientation and Right Wing Authoritarianism, which 
control for one another within the same model. Figure \@ref(fig:plotAllPolRel2)
in the main text shows the same results, but from models fitted to the reduced
dataset with pre-registered exclusions. Points represent posterior medians, line 
ranges represent 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

```{r plotAllPolRel1, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPolRel1Caption)", warning=FALSE}
tar_read(plotAllPolRel1)
```

\newpage

(ref:plotSliders1Caption) *Boxplots showing the distribution of responses to each
self-report question about the reasons for participants' behaviour in the games.*
Boxplots represent medians and interquartile ranges.

```{r plotSliders1, fig.width=7, fig.height=4, fig.cap="(ref:plotSliders1Caption)", warning=FALSE}
tar_read(plotSliders1)
```

\newpage

(ref:plotSliders2Caption) *Boxplots showing the distribution of responses to each
self-report question about the reasons for participants' behaviour in the games,
presented as deviations from participants' average rating across all questions.*
Boxplots represent medians and interquartile ranges.

```{r plotSliders2, fig.width=7, fig.height=4, fig.cap="(ref:plotSliders2Caption)", warning=FALSE}
tar_read(plotSliders2)
```

\newpage

(ref:plotAllSliders1Caption) *Posterior slopes from models including 
self-reported strategy usage, fitted to the full dataset without pre-registered 
exclusions.* Each row represents a separate model. Highlighted estimates 
represent combinations where the self-report slider matched the behavioural 
strategy. Figure \@ref(fig:plotAllSliders2) in the main text shows the same 
results, but from models fitted to the reduced dataset with pre-registered 
exclusions. Points represent posterior medians, line ranges represent 95% 
credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.

```{r plotAllSliders1, fig.width=7.5, fig.height=4, fig.cap="(ref:plotAllSliders1Caption)", warning=FALSE}
tar_read(plotAllSliders1)
```

\newpage

(ref:plotSplineCaption) *The relationships between strategy frequencies and the
overall costs of strategies across all twelve punishment decisions, in both
countries.* Each point is a unique strategy that appears in our dataset at
least once (for ease of presentation, the "never punish" strategy is excluded).
Lines and shaded areas represent posterior predictions from splines fitted to
the trend in each country separately.

```{r plotSpline, fig.width=6, fig.height=3.5, fig.cap="(ref:plotSplineCaption)", warning=FALSE}
tar_read(plotSpline)
```

\newpage

(ref:plotSampleUKCaption) *Sample characteristics in the United Kingdom.*

```{r plotSampleUK, fig.width=7, fig.height=5, fig.cap="(ref:plotSampleUKCaption)", warning=FALSE}
tar_read(plotSample_UK)
```

\newpage

(ref:plotSampleUSCaption) *Sample characteristics in the United States.*

```{r plotSampleUS, fig.width=7, fig.height=5, fig.cap="(ref:plotSampleUSCaption)", warning=FALSE}
tar_read(plotSample_US)
```

\newpage

(ref:plotModel5Caption) *Posterior estimates of the probabilities of following
different punishment strategies from the Bayesian latent state model that
estimates the implementation error rate as a free parameter.* The model
estimated the implementation error rate to be `r printnum(median(error))`
(95% CI [`r printnum(quantile(error, 0.025))`
`r printnum(quantile(error, 0.975))`]). Points represent posterior medians, line 
ranges represent 50% and 95% credible intervals. AI = advantageous inequity, DI
= disadvantageous inequity.

```{r plotModel5, fig.width=7, fig.height=5, fig.cap="(ref:plotModel5Caption)", warning=FALSE}
tar_read(plot5)
```

\newpage

(ref:plotTraceCaption) *MCMC trace plots for parameter values from the Bayesian
latent state model fitted to data with exclusions.* These trace plots suggest
that the different chains mixed well and the model converged normally.

```{r plotTrace, fig.width=7, fig.height=6, fig.cap="(ref:plotTraceCaption)", warning=FALSE}
tar_read(plotTrace1.2)
```

\newpage

(ref:plotSimCaption) *Results of Bayesian latent state model fitted to simulated
data (n = 100) with known strategy frequencies in the population.* Blue points
represent known strategy frequencies, grey densities represent posterior
estimates of strategy frequencies. AI = advantageous inequity, DI = 
disadvantageous inequity.

```{r plotSim, fig.width=6, fig.height=5, fig.cap="(ref:plotSimCaption)", warning=FALSE}
tar_read(plotSim1)
```

\newpage

## Supplementary Tables

(ref:tablePatternsCaption) Counts and proportions of the 25 most common patterns
of punitive behaviour across all twelve decisions, split by country. *Binary 
strings represent punishment (1) or no punishment (0) in each decision, aligning 
with the order of game decision columns in Table \@ref(tab:tableStrategies).*

```{r tablePatterns}
apa_table(
  tar_read(tablePatterns),
  escape = FALSE,
  booktabs = TRUE,
  align = c(rep("l", 2), rep("c", 4)),
  placement = "H",
  col_spanners = list(
    "\\thead{United Kingdom \\\\ (N = 1014)}" = 3:4,
    "\\thead{United States \\\\ (N = 996)}" = 5:6
    ),
  caption = "(ref:tablePatternsCaption)",
  font_size = "footnotesize"
)
```

\newpage

(ref:tableSliderWordingsCaption) Wordings for 11 self-report slider questions
asking participants to report the reasons for their behaviour in the six games.
*Participants were prompted with the following text: "We would now like you to 
answer a few questions about your main motivation in the games. Please answer 
truthfully - there is no right or wrong answer and your first answer is probably 
best. Please rate the extent to which the following statements apply to your 
decisions to reduce or not to reduce other players' bonuses in the games."*

```{r tableSliderWordings}
apa_table(
  tar_read(tableSliderWordings),
  caption = "(ref:tableSliderWordingsCaption)",
  booktabs = TRUE,
  placement = "H",
  font_size = "footnotesize"
)
```

\newpage

(ref:tableSurveyWordingsCaption) Wordings for survey questions in the study.

```{r tableSurveyWordings}
apa_table(
  tar_read(tableSurveyWordings),
  landscape = TRUE,
  longtable = TRUE,
  align = c("p{6cm}", "p{11cm}", "p{3cm}"),
  caption = "(ref:tableSurveyWordingsCaption)",
  booktabs = TRUE,
  placement = "H",
  font_size = "scriptsize"
)
```

\newpage

(ref:tableCompCaption) Proportions of correct answers to comprehension 
questions for all six economic games, split by country.

```{r tableComp}
apa_table(
  tar_read(tableComp),
  caption = "(ref:tableCompCaption)",
  booktabs = TRUE,
  placement = "H"
)
```