### 2025-08-01 quarto/appendix.qmd
Update manuscript and submit

# Supplementary Materials

## Supplementary Figures

::: {#suppfig-model1a suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot1.1)
```

Posterior estimates of the probabilities of following different punishment 
strategies from the Bayesian latent state model fitted to the full dataset 
without pre-registered exclusions. The model assumes an implementation error 
rate of 5%. @fig-model1b in the main text shows the same result, but from a 
model fitted to the reduced dataset with pre-registered exclusions. Points 
represent posterior medians, line ranges represent 50% and 95% credible 
intervals. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-allDems2 suppfig-pos="H"}

```{r, fig.height=3, fig.width=7.5, warning=FALSE}
tar_read(plotAllDems2)
```

Posterior slopes from models including age, socio-economic status, gender, and 
student status, fitted to the subsetted dataset with pre-registered exclusions.
Each row represents a separate model. Points represent posterior medians, line 
ranges represent 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

:::

\newpage

::: {#suppfig-allDems1 suppfig-pos="H"}

```{r, fig.height=3, fig.width=7.5, warning=FALSE}
tar_read(plotAllDems1)
```

Posterior slopes from models including age, socio-economic status, gender, and 
student status, fitted to the full dataset without pre-registered exclusions. 
Each row represents a separate model. Points represent posterior medians, line 
ranges represent 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

:::

\newpage

::: {#suppfig-allPers1 suppfig-pos="H"}

```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
tar_read(plotAllPers1)
```

Posterior slopes from models including Big-6 personality dimensions and Social 
Value Orientation, fitted to the full dataset without pre-registered exclusions. 
Each row represents a separate model. @fig-allPers2 in the main text shows the 
same results, but from models fitted to the reduced dataset with pre-registered 
exclusions. Points represent posterior medians, line ranges represent 95% 
credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-allPolRel1 suppfig-pos="H"}

```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
tar_read(plotAllPolRel1)
```

Posterior slopes from models including political ideology, views about social 
inequality, and religiosity, fitted to the full dataset without pre-registered 
exclusions. Each row represents a separate model aside from Social Dominance 
Orientation and Right Wing Authoritarianism, which control for one another 
within the same model. @fig-allPolRel2 in the main text shows the same results,
but from models fitted to the reduced dataset with pre-registered exclusions. 
Points represent posterior medians, line ranges represent 95% credible 
intervals. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-sliders1 suppfig-pos="H"}

```{r, fig.height=4, fig.width=7, warning=FALSE}
tar_read(plotSliders1)
```

Boxplots showing the distribution of responses to each self-report question 
about the reasons for participants' behaviour in the games. Boxplots represent 
medians and interquartile ranges.

:::

\newpage

::: {#suppfig-sliders2 suppfig-pos="H"}

```{r, fig.height=4, fig.width=7, warning=FALSE}
tar_read(plotSliders2)
```

Boxplots showing the distribution of responses to each self-report question 
about the reasons for participants' behaviour in the games, presented as 
deviations from participants' average rating across all questions. Boxplots 
represent medians and interquartile ranges.

:::

\newpage

::: {#suppfig-allSliders1 suppfig-pos="H"}

```{r, fig.height=4, fig.width=7.5, warning=FALSE}
tar_read(plotAllSliders1)
```

Posterior slopes from models including self-reported strategy usage, fitted to 
the full dataset without pre-registered exclusions. Each row represents a 
separate model. Highlighted estimates represent combinations where the 
self-report slider matched the behavioural strategy. @fig-allSliders2 in the 
main text shows the same results, but from models fitted to the reduced dataset
with pre-registered exclusions. Points represent posterior medians, line ranges
represent 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

:::

\newpage

::: {#suppfig-spline suppfig-pos="H"}

```{r, fig.height=3.5, fig.width=6, warning=FALSE}
tar_read(plotSpline)
```

The relationships between strategy frequencies and the overall costs of 
strategies across all twelve punishment decisions, in both countries. Each point
is a unique strategy that appears in our dataset at least once (for ease of 
presentation, the "never punish" strategy is excluded). Lines and shaded areas
represent posterior predictions from splines fitted to the trend in each country 
separately.

:::

\newpage

::: {#suppfig-sampleUK suppfig-pos="H"}

```{r, fig.height=5, fig.width=7, warning=FALSE}
tar_read(plotSample_UK)
```

Sample characteristics in the United Kingdom.

:::

\newpage

::: {#suppfig-sampleUS suppfig-pos="H"}

```{r, fig.height=5, fig.width=7, warning=FALSE}
tar_read(plotSample_US)
```

Sample characteristics in the United States.

:::

\newpage

::: {#suppfig-model5 suppfig-pos="H"}

```{r, fig.height=5, fig.width=7, warning=FALSE}
tar_read(plot5)
```

Posterior estimates of the probabilities of following different punishment 
strategies from the Bayesian latent state model that estimates the 
implementation error rate as a free parameter. The model estimated the 
implementation error rate to be `r papaja::printnum(median(error))`
(95% CI [`r papaja::printnum(quantile(error, 0.025))`
`r papaja::printnum(quantile(error, 0.975))`]). Points represent posterior 
medians, line ranges represent 50% and 95% credible intervals. AI = advantageous 
inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-trace suppfig-pos="H"}

```{r, fig.height=6, fig.width=7, warning=FALSE}
tar_read(plotTrace1.2)
```

MCMC trace plots for parameter values from the Bayesian latent state model 
fitted to data with exclusions. These trace plots suggest that the different
chains mixed well and the model converged normally.

:::

\newpage

::: {#suppfig-sim suppfig-pos="H"}

```{r, fig.height=5, fig.width=6, warning=FALSE}
tar_read(plotSim1)
```

Results of Bayesian latent state model fitted to simulated data (n = 100) with 
known strategy frequencies in the population. Blue points represent known 
strategy frequencies, grey densities represent posterior estimates of strategy 
frequencies. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

## Supplementary Tables

::: {#supptbl-patterns supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tablePatterns) %>%
  flextable() %>%
   set_header_labels(
    Pattern = "Pattern",
    Explanation = "Explanation",
    a = "N",
    b = "Prop",
    c = "N",
    d = "Prop"
  ) %>%
  add_header_row(
    values = c(
      "",
      "United Kingdom\n(N = 1014)",
      "United States\n(N = 996)"
    ),
    colwidths = c(2, 2, 2)
  ) %>%
  theme_apa() %>%
  line_spacing(part = "all") %>%
  align(i = NULL, j = 1:2, align = "left", part = "all") %>%
  autofit() %>%
  fontsize(size = 9.5, part = "all")
```

Counts and proportions of the 25 most common patterns of punitive behaviour 
across all twelve decisions, split by country. Binary strings represent 
punishment (1) or no punishment (0) in each decision, aligning with the order of
game decision columns in @tbl-strategies.

:::

\newpage

::: {#supptbl-sliderWordings supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tableSliderWordings) %>%
  flextable() %>%
  theme_apa() %>%
  autofit() %>%
  align(i = NULL, j = 2, align = "left", part = "all") %>%
  line_spacing(part = "all")
```

Wordings for 11 self-report slider questions asking participants to report the 
reasons for their behaviour in the six games. Participants were prompted with 
the following text: "We would now like you to answer a few questions about your
main motivation in the games. Please answer truthfully - there is no right or 
wrong answer and your first answer is probably best. Please rate the extent to
which the following statements apply to your decisions to reduce or not to
reduce other players' bonuses in the games."

:::

\newpage

::: {#supptbl-surveyWordings supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tableSurveyWordings) %>%
  flextable() %>%
  theme_apa() %>%
  autofit() %>%
  align(align = "left", part = "all") %>%
  line_spacing(part = "all")
```

Wordings for survey questions in the study.

:::

\newpage

::: {#supptbl-comp supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tableComp) %>%
  flextable() %>%
  theme_apa() %>%
  autofit() %>%
  align(j = 1, align = "left", part = "all")
```

Proportions of correct answers to comprehension questions for all six economic 
games, split by country.

:::