### 2025-08-01 quarto/manuscript.qmd

---
title: "Why Do People Punish? Evidence for a Range of Strategic Concerns"
shorttitle: "Punishment strategies"
author:
  - name: Scott Claessens
    corresponding: true
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        city: Canterbury
        country: United Kingdom
      - name: University of Auckland
        department: School of Psychology
        city: Auckland
        country: New Zealand
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Quentin D Atkinson
    corresponding: false
    orcid: 0000-0002-8499-7535
    affiliations:
      - name: University of Auckland
        department: School of Psychology
        city: Auckland
        country: New Zealand
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - writing
      - editing
  - name: Nichola J Raihani
    corresponding: false
    orcid: 0000-0003-2339-9889
    affiliations:
      - name: University of Auckland
        department: School of Psychology
        city: Auckland
        country: New Zealand
      - name: University College London
        department: Department of Experimental Psychology
        city: London
        country: United Kingdom
    role:
      - conceptualization
      - funding acquisition
      - investigation
      - methodology
      - supervision
      - writing
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: "This manuscript was published as a pre-print on PsyArXiv under a CC BY 4.0 license: <https://doi.org/10.31234/osf.io/ys6rm>"
    data-sharing: Word count (excluding abstract, figures, tables, and references) = 6065 words.
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "Costly punishment is thought to be a key mechanism sustaining human cooperation. However, the motives for punitive behaviour remain unclear. Although often assumed to be motivated by a desire to convert cheats into cooperators, punishment is also consistent with other functions, such as levelling payoffs or improving one’s relative position. We used six economic games to tease apart different motives for punishment and to explore whether different punishment strategies were associated with personality variables, political ideology, and religiosity. We used representative samples from the United Kingdom and the United States (N = 2010) to estimate the frequency of different punishment strategies in the population. The most common strategy was to never punish: prosocial individuals and those scoring high in honesty-humility were among the least punitive in our sample, while religious individuals and those scoring high for right-wing authoritarianism and social dominance orientation were more punitive. For people who did punish, strategy use was more consistent with egalitarian motives than behaviour-change motives. Nevertheless, different punishment strategies were also associated with personality, social preferences, political ideology, and religiosity. Self-reports of behaviour in the games suggested that people have some insight into their punishment strategy. These findings highlight the multipurpose nature of human punishment and show how the different motives underpinning punishment decisions are linked with core character traits."
keywords: [punishment, cooperation, economic games]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-docx:
    fig-dpi: 300
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
csl: nature.csl
---

```{r}
library(flextable)
library(knitr)
library(targets)
library(tidyverse)
```

Humans cooperate on a scale that is unparalleled in the animal kingdom. One 
mechanism thought to sustain this level of cooperation is costly punishment, 
whereby individuals harm others at a personal cost [@CluttonBrock1995], 
ostensibly encouraging cooperative behaviour from the target or bystanders 
[@DosSantos2011; @DosSantos2013; @Raihani2015]. Punishment offers a route to 
maintaining or increasing cooperation by changing the payoff structure of social 
interactions such that it no longer pays to cheat or exploit social partners 
[@CluttonBrock1995; @Ostrom1990]. Yet, despite its theoretical importance, the 
question of why people choose to punish others is still hotly contested 
[@Raihani2019]. In this study, we use a battery of economic games to disentangle
the different motives underpinning punishment and explore how these motives vary 
across individuals.

In humans, punishment is often studied in laboratory settings using economic 
games [@Balliet2011; @Balliet2013; @Chaudhuri2011; @Dreber2008; @Fehr2000; 
@Fehr2002; @Henrich2006; @Nikiforakis2008a; @Raihani2012a; @Raihani2019]. In 
some of these games, participants are given a sum of money that they can use to 
invest in collective action or to help others; conversely, they can 'cheat' by 
keeping the money for themselves or by exploiting the contributions of others. 
In other games, participants are endowed with sums of money and can then 'cheat'
by destroying a partner's payoffs or stealing from them [@Abbink2009; 
@Raihani2012b]. Punishment is introduced into such games by giving participants 
the option to pay a small 'fee' to impose a greater 'fine' on their co-players. 
Several lines of experimental evidence indicate that people use this punishment 
option [@Henrich2006], that they enjoy punishing [@DeQuervain2004], and that 
they frequently, though not always [@Herrmann2008], punish those who cheat,
free-ride, or steal [@Fehr2000; @Fehr2002].

Evidence from these experiments suggests that the threat of costly punishment 
plays an important role in promoting human cooperation. People tend to cooperate 
more in games where punishment is possible compared to those where it is not
[@Balliet2011; @Balliet2013; @Raihani2019]. Similarly, people make higher offers 
in the Ultimatum Game (where punishment is possible) compared to the 
structurally-similar Dictator Game (where it is not) [@Camerer2003]. This 
cooperation-enhancing effect of punishment has also been observed across 
societies [@Balliet2013], leading some to suggest that costly punishment has 
played a key role in the cultural evolution of cooperation in humans
[@Bowles2013; @Boyd2003; @Chudek2011; @Henrich2017].

Nevertheless, it remains unclear whether individuals playing economic games use
punishment as a behaviour-change tool to enforce cooperation or to achieve other 
ends. Some have argued that punishment is primarily used to shape behaviour, 
either to deter personal harm [@Delton2017; @Dreber2008; @DosSantos2013] or to 
uphold normative standards of cooperative behaviour [@Boyd2003; @Chudek2011; 
@Fehr2018; @Mathew2011; @Mathew2014; @Richerson2016]. But while the _threat_ of 
punishment can have a cooperation-enhancing effect, the _enactment_ of this 
punishment does not consistently deter targets from cheating in the future 
[@Raihani2019]. This calls into question whether punishment primarily operates 
as a behaviour-change tool or whether it is used to achieve other goals.

Beyond behaviour shaping concerns, there are other reasons why people punish in 
economic games. Punishers might be motivated by a desire for retribution rather 
than deterrence, punishing in proportion to the amount of harm that was 
personally caused [@Carlsmith2002]. Punishment might be driven by concerns about 
relative payoffs, such as disadvantageous inequity aversion (i.e., avoiding 
having less than others) [@Raihani2012b; @Raihani2019] and/or general 
egalitarian preferences (i.e., wanting all participants to receive the same 
payoffs) [@Dawes2007]. Such concerns about relative payoffs may be activated 
when participants earn less than cheats in economic games or when there are 
income disparities in these settings. People might also use punishment for 
competitive purposes, seeking advantageous inequity for themselves (i.e., having 
more than others) and/or improving their relative position [@Raihani2019].

Common economic game designs have been unable to tease apart these different 
motives for punishment because participants who interact with cheats in these 
games experience both losses _and_ lower relative payoffs. The typical 1:3 
fee-fine ratio of punishment in economic games compounds this issue. With 
this setup, people can simultaneously use punishment to reciprocate losses, to 
deter others from cheating, and to reduce or reverse disparities in payoffs 
between themselves and targets. To add to this complexity, it is evident that 
people use punishment in seemingly disparate ways: punishing when no behaviour 
change is possible, such as in one-shot games [@Bone2015; @Henrich2006;
@Raihani2012b; @Walker2004], on the very last round of repeated games 
[@Barclay2016], or in games where the target never learns about the punishment 
[@Crockett2014]; punishing those who did not cheat or who over-contributed to 
collective action (sometimes called 'antisocial punishment') [@Herrmann2008; 
@Sylwester2013]; punishing in scenarios where they were not personally harmed 
(third-party punishment) [@Fehr2004]; and punishing in scenarios where 
disparities in payoffs did not arise from participants' actions [@Dawes2007; 
@Johnson2009; @Fowler2005].

The general conclusion from this research is that there is no one unifying 
function of costly punishment in humans. Instead, punishment should be thought 
of as a flexible behavioural tool that serves a variety of functions that are 
not mutually exclusive [@Raihani2019]. Due to its multipurpose nature, we should 
therefore expect variation in punishment strategies in the population, much like 
the observed variation in social learning strategies [@Molleman2014]. Some 
individuals may use punishment as a behaviour shaping tool, for example, while 
others may use it to reduce or reverse payoff differentials.

This insight raises several underexplored questions. First, which punishment 
strategies are more frequent in human populations? Second, what traits predict 
adherence to a particular strategy? Previous work has reported that personality 
is related to cooperative behaviour [@Thielmann2020] and demographics, political 
ideology, and religiosity are related to punitive behaviour [@Hofmann2018], but 
no research has related these variables to specific punishment strategies. 
Third, do people have insight into their own punishment strategy? Previous work 
has argued that people are often unaware of the underlying function of their 
punitive behaviour, yet they feel compelled to enact it anyway
[@Carlsmith2002; @Nisbett1977].

Here, we aim to delineate nine possible punishment strategies by asking 
whether people punish in a manner consistent with a specific strategy and, if 
so, what other characteristics (personality, social preferences, political 
orientation, religious views) predict the use of different punishment 
strategies. @tbl-strategies summarises the potential functions for costly 
punishment in the economic games that we considered, and the behavioural 
strategies they predict. Note that @tbl-strategies is not an exhaustive list of 
all possible punishment strategies: we do not include reputational functions of 
punishment in this table, such as signalling trustworthiness [@Barclay2006; 
@Batistoni2022; @Jordan2016; @Jordan2017; @Raihani2015], because our focus is on 
punishment strategies in anonymous economic games without reputational 
incentives (but see ref [@Jordan2020]).

Building on previous designs that have used one-shot economic games to explore 
behaviour-shaping and egalitarian motives for punishment [@Bone2015; 
@Crockett2014; @Deutchman2021; @Marczyk2017; @Raihani2012b], we employ a suite 
of one-shot economic games where individuals are given the opportunity to punish 
targets at a personal cost (@fig-games). In each game, targets either steal from 
another individual or do nothing. Representative samples of participants from 
the United Kingdom (*n* = `r sum(tar_read(dExc)$Country == "United Kingdom")`) 
and the United States (*n* = `r sum(tar_read(dExc)$Country == "United States")`) 
completed all six games on the online platform Prolific. We designed the suite 
of games to tease apart the proposed punishment strategies in @tbl-strategies, 
such that each strategy predicts a different pattern of behaviour across all the 
games (see Methods for more detail about the six games). We use the resulting 
behavioural patterns to discern which punishment strategy participants are 
employing. We then combine these behavioural patterns with data on demographics, 
personality, social preferences, political ideology, religiosity, and self-reported strategy usage.

![Visual summary of the six economic games.](images/games.png){
  #fig-games
  width="80%"
  apa-note="In the games, Player 2 either steals £0.20 from Player 1 (the focal player) or does nothing. Player 1 is then given the option to punish by paying a certain amount of money to remove money from Player 2 (this money is destroyed). The six games are variants on this general setup, creating situations where (A) Player 2 is still worse off after stealing, (B) Player 2 creates equality by stealing, (C) the computer 'decides' whether Player 2 steals, (D) the fee-fine ratio is 1:1, (E) Player 2 is better off after stealing, and (F) Player 2 steals instead from a third-party."
  }

::: {.landscape}

```{r}
#| label: tbl-strategies
#| tbl-cap: Summary of the different functions for punishment and the behavioural strategies they predict.
#| apa-note: Games A-F are the games employed in the current study (see Methods for more details). In each of the six games, participants are given the opportunity to punish players who 'steal' and those who do not, meaning that participants make twelve punishment decisions in total. Each behavioural strategy implies a unique pattern of punishment across all decisions (see Methods for detailed explanation of strategies). Ticks reflect decisions to punish, crosses reflect decisions to not punish. In column headers, payoffs at the first stage (above) and the second stage (below) are denoted as P1-P2 (or P2-P3 [P1] for Game F) where participants take the role of P1 and P2 is the target of punishment. AI = advantageous inequity, DI = disadvantageous inequity.

tar_read(tableStrategies) %>%
  flextable() %>%
  set_header_labels(
    Function = "Function",
    `Behavioural strategy` = "Behavioural strategy",
    a = "Steal \n 50-30",
    b = "No steal \n 70-10",
    c = "Steal \n 50-50",
    d = "No steal \n 70-30",
    e = "Steal \n 50-50",
    f = "No steal \n 70-30",
    g = "Steal \n 50-50",
    h = "No steal \n 70-30",
    i = "Steal \n 50-70",
    j = "No steal \n 70-50",
    k = "Steal \n 50-90 [100]",
    l = "No steal \n 70-70 [100]"
  ) %>%
  add_header_row(
    values = c(
      "",
      "Game A \n (AI) \n 70-10",
      "Game B \n (Equal) \n 70-30",
      "Game C \n (Computer) \n 70-30",
      "Game D \n (1:1 Fee Fine) \n 70-30",
      "Game E \n (DI) \n 70-50",
      "Game F \n (Third Party) \n 70-70 [100]"
    ),
    colwidths = rep(2, 7)
  ) %>%
  theme_apa() %>%
  width(width = c(1, 2, rep(0.6, 12))) %>%
  fontsize(size = 9, part = "all") %>%
  line_spacing(part = "all") %>%
  align(i = NULL, j = 1:2, align = "left", part = "all")
```

:::

# Results

```{r MLMs}
mlm1 <- coef(summary(tar_read(mlm1)))
mlm2 <- coef(summary(tar_read(mlm2)))
```

The overall pattern of punitive behaviour in the six economic games was very 
similar across both countries (@fig-decisions). Participants were generally more 
likely to punish targets who stole compared to targets who did not steal 
(multilevel logistic regression; _b_ = 
`r papaja::printnum(mlm1["otherTook","Estimate"])`, standard error = 
`r papaja::printnum(mlm1["otherTook","Std. Error"])`, _p_ 
`r papaja::printp(mlm1["otherTook","Pr(>|z|)"])`). Participants were also more 
likely to punish when targets' stealing behaviour generated inequalities, 
specifically in Games E and F (_b_ = 
`r papaja::printnum(mlm2["DI","Estimate"])`, SE = 
`r papaja::printnum(mlm2["DI","Std. Error"])`, 
_p_ `r papaja::printp(mlm2["DI","Pr(>|z|)"])`).

```{r, fig.height=5, fig.width=8}
#| label: fig-decisions
#| fig-cap: Overall pattern of punitive behaviour across all six economic games, split by country.
#| fig-align: center
#| apa-note: AI = advantageous inequity, DI = disadvantageous inequity.

tar_read(plotDecisions)
```

<br>

We classified participants into a particular strategy if their behaviour across 
all twelve decisions matched our behavioural predictions shown in 
@tbl-strategies exactly. @tbl-strategycounts shows the proportion of 
participants following each strategy, with N/A used to represent participants 
who did not fit exactly into any particular strategy type. Overall, 
`r papaja::printnum(mean(tar_read(dExc)$strategy != "N/A")*100, digits = 0)`%
of our participants could be classified exactly into one of the strategies. The
most common strategy in both countries was to never punish across any of the
games. The next most common strategies were those that care about
minimising payoff differences (avoid disadvantageous inequity, egalitarian). 
Less common were the behaviour shaping strategies (deterrent, norm-enforcing),
the retributive strategy, and the competitive strategies (seek advantageous 
inequity, competitive). Although participants often punished targets who did not 
steal in the six games (@fig-decisions), no participants followed the antisocial
strategy by exclusively punishing targets who did not steal across _all_ games.

```{r}
#| label: tbl-strategycounts
#| tbl-cap: Counts and proportions of participants following each punishment strategy exactly, split by country.
#| apa-note: N/A implies that participants were unable to be classified exactly into any of the punishment strategies. AI = advantageous inequity, DI = disadvantageous inequity.

tar_read(tableStrategyCounts) %>%
  flextable() %>%
  set_header_labels(
    Strategy = "Strategy",
    a = "N",
    b = "Prop",
    c = "N",
    d = "Prop"
  ) %>%
  add_header_row(
    values = c(
      "",
      "United Kingdom\n(N = 1014)",
      "United States\n(N = 996)"
    ),
    colwidths = c(1, 2, 2)
  ) %>%
  autofit() %>%
  theme_apa() %>%
  line_spacing(part = "all") %>%
  align(i = NULL, j = 1, align = "left", part = "all")
```

<br>

To further investigate the strategies that participants were following, we 
examined the most common patterns of punitive behaviour across all twelve 
decisions. @supptbl-patterns shows the proportion of participants following the
25 most common behavioural patterns, including, where appropriate, the 
predetermined strategies from @tbl-strategies. In both countries, a common 
pattern of behaviour not captured by any of the strategies was punishing only 
when the target stole in the third-party game (Game F). Punishment in this game
is consistent with an egalitarian motive, as stealing produces unequal outcomes,
but third-party punishment here is also consistent with norm-enforcing and 
competitive motives (see @tbl-strategies). Other common behavioural patterns not 
captured by our strategies included punishing whenever the target stole across 
all games and always punishing in every game irrespective of the targets' 
behaviour.

While it is useful to look at exact patterns of behaviour, participants may not 
have implemented their chosen punishment strategy with exact precision.
In reality, strategies may have been implemented probabilistically for each 
punishment decision. There is also the possibility of implementation errors, 
whereby participants occasionally "slip up" and make decisions that are 
incongruent with a particular strategy. This may explain why some participants 
were unable to be classified exactly into a single punishment strategy.

To deal with this complexity and include all observed data in our frequency 
estimates, we fitted a Bayesian latent state model to the data. This model 
assumes that the nine strategies in @tbl-strategies (plus a "random choice" 
strategy that chooses randomly for each decision) are the only latent strategies 
and that these are instantiated into observed behaviour according to the logic 
in @tbl-strategies with some probability of implementation error (i.e., an 
intention to punish is implemented as non-punishment and vice versa). Averaging 
over all strategies and incorporating the possibility of implementation errors,
the model estimates the probability of participants following any particular 
strategy, given the observed data.

The posterior estimates from the model are presented in @fig-model1b. The 
posterior probabilities for each strategy did not differ between the two 
countries. In both countries, the never punish strategy had the highest 
probability, followed by the egalitarian strategy. The norm-enforcing and seek 
advantageous inequity strategies were the next most likely, with higher 
posterior estimates than the competitive and antisocial strategies. None of the 
other strategies differed in their posterior estimates. The same general pattern 
emerged when we analysed the full dataset without pre-registered exclusions 
(@suppfig-model1a).

```{r, fig.height=5, fig.width=7}
#| label: fig-model1b
#| fig-cap: Posterior estimates of the probabilities of following different punishment strategies from the Bayesian latent state model.
#| fig-align: center
#| apa-note: The model assumes an implementation error rate of 5%. Points represent posterior medians, line ranges represent 50% and 95% credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.

tar_read(plot1.2)
```

<br>

Next, we explored which traits predicted adherence to different punishment 
strategies. To answer this question, we included variables capturing 
demographics, personality, social preferences, political views, and religious 
views as predictors in our Bayesian latent state model. We included each 
variable in a separate model, predicting all ten punishment strategies 
(the nine from @tbl-strategies, plus the 'random choice' strategy) 
simultaneously.

```{r sesSlope}
ses <- tar_read(post2.2_SES)$beta[,2,10]
```

Demographic variables tended to be unrelated to strategy usage: age and gender
did not predict adherence to a particular punishment strategy (Supplementary 
Figures -@suppfig-allDems2 and -@suppfig-allDems1). In the United States, the 
never punish strategy was slightly more common among participants lower in 
socio-economic status (median posterior slope = 
`r papaja::printnum(median(ses))`, 95% CI
[`r papaja::printnum(quantile(ses, 0.025))` 
`r papaja::printnum(quantile(ses, 0.975))`]) but this effect was small.

Conversely, personality and social preferences were linked to variation in 
punishment strategies. When including the Big-6 personality dimensions and 
Social Value Orientation (SVO) in the model, we found associations with the 
egalitarian, never punish, and random choice strategies, with small-to-medium 
effect sizes (@fig-allPers2). More prosocial participants (those with higher SVO 
scores) were more likely to follow the egalitarian and the never punish 
strategies, while those with lower SVO scores were more likely to enact the 
random choice strategy. The personality dimensions of honesty-humility and 
openness to experience were both positively associated with following the never 
punish strategy, while extraversion negatively predicted this strategy. The 
effects were mostly similar across countries, but occasionally differed: for 
example, in the United States, but not in the United Kingdom, honesty-humility 
was positively associated with following the egalitarian strategy and negatively 
associated with following the random choice strategy. Overall, the same pattern
of results emerged when analysing the full dataset without exclusions 
(@suppfig-allPers1).

```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
#| label: fig-allPers2
#| fig-cap: Posterior slopes from Bayesian latent state models including Big-6 personality dimensions and Social Value Orientation.
#| fig-align: center
#| apa-note: Each row represents a separate model. All models assume an implementation error rate of 5%. Points represent posterior medians, line ranges represent 95% credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.

tar_read(plotAllPers2)
```

<br>

Political and religious variables were also associated with punishment strategy 
(@fig-allPolRel2). These effects were small-to-medium in size and tended to be 
more pronounced in the United States. Controlling for Social Dominance 
Orientation, American participants higher in Right Wing Authoritarianism were 
more likely to follow the strategies avoiding disadvantageous inequity and 
seeking advantageous inequity. Participants who stated that they would like to 
"bring those below them [on the socio-economic status ladder] up a peg" were 
more likely to follow the egalitarian strategy, while American participants 
higher in Social Dominance Orientation, Right Wing Authoritarianism, and 
believing that God controls events in the world were less likely to follow the 
egalitarian strategy. In general, religious and conservative participants were 
less likely to follow the never punish strategy. This general pattern of results 
was replicated with the full dataset (@suppfig-allPolRel1).

```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
#| label: fig-allPolRel2
#| fig-cap: Posterior slopes from Bayesian latent state models including political ideology, views about social inequality, and religiosity.
#| fig-align: center
#| apa-note: Each row represents a separate model aside from Social Dominance Orientation and Right Wing Authoritarianism, which control for one another within the same model. All models assume an implementation error rate of 5%. Points represent posterior medians, line ranges represent 95% credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.

tar_read(plotAllPolRel2)
```

<br>

Finally, we asked whether participants had insight into their own punishment 
strategy. In other words, could participants self-report the strategy that they 
were following during the games? To answer this question, we included 
participants' responses to post-game questions about their strategy as 
predictors in the model. As before, each predictor was included in a separate 
model, predicting all ten strategies simultaneously.

In general, we found that self-reported strategy usage was positively associated
with the behavioural strategy that participants employed, with effects ranging 
from small to large in size (see Supplementary Figures -@suppfig-sliders1 and 
-@suppfig-sliders2 for the distribution of responses to self-report questions).
@fig-allSliders2 shows the relationships between self-report questions and the 
different punishment strategies, highlighting the combinations where the 
question matched the behavioural strategy. We found positive relationships 
between the self-report questions and strategy usage for the norm-enforcing, 
egalitarian, seek advantageous inequity, never punish, and random choice 
strategies. The 95% credible intervals for other estimates included zero, though
these estimates often trended in a positive direction. The same pattern of
results was found when analysing the full dataset without exclusions
(@suppfig-allSliders1).

```{r, fig.height=4, fig.width=7.5, warning=FALSE}
#| label: fig-allSliders2
#| fig-cap: Posterior slopes from models including self-reported strategy usage.
#| fig-align: center
#| apa-note: Each row represents a separate model. All models assume an implementation error rate of 5%. Highlighted estimates represent combinations where the self-report question matched the behavioural strategy. Points represent posterior medians, line ranges represent 95% credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.

tar_read(plotAllSliders2)
```

<br>

# Discussion

Using a suite of economic games measuring punishment in different situations, 
we have shown that punishment does not serve just one function, but instead is 
a flexible tool that can be and is used for different purposes [@Raihani2019]. 
While some use punishment to enforce norms of cooperation, others use it to 
reduce or even create inequality between individuals. We found that people's 
punishment strategy can, to some extent, be predicted by individual differences 
in personality, social preferences, and political and religious views. Moreover, 
contrary to the view that people are often unable to articulate the reasons for 
their punitive behaviour [@Carlsmith2002], people seem to have some degree of 
insight into the strategy they are using. Despite small differences, these 
general patterns replicated in samples from both the United Kingdom and the 
United States, providing further confidence in the results.

Among the punitive strategies, the most common were particularly sensitive to 
inequality in payoffs, either from a self-referential perspective (i.e., avoid 
disadvantageous inequity) or more generally (i.e., egalitarian). This is in 
line with previous studies which have highlighted inequity aversion as an 
important motivation for punishment in economic games [@Bone2015; @Bone2016; 
@Dawes2007; @Deutchman2021; @Johnson2009; @Raihani2012b]. Personality, social 
preferences, and political and religious views were related to these strategies.
Traits associated with other-regarding concern, such as SVO and 
honesty-humility, predicted following the egalitarian strategy. By contrast,
religious and conservative individuals, including individuals higher in SDO and
RWA, were less likely to follow the egalitarian strategy and more likely to 
follow the avoid disadvantageous inequity strategy, especially in the United 
States. Authoritarian participants in the United States were also more likely to 
actively seek advantageous inequity with their punishment. These findings 
provide insight into the motives that may have driven higher levels of peer 
punishment among conservatives in previous work [@Hofmann2018].

Behaviour shaping strategies, such as deterrence and norm-enforcement, were less 
common than strategies sensitive to inequality in our set of games. This was 
reflected both in participants’ elicited punishment behaviour (@fig-model1b)
and in their self-reports of their own strategy (Supplementary Figures 
-@suppfig-sliders1 and -@suppfig-sliders2). Although our design did not 
explicitly allow for behaviour shaping as the interactions were all one-shot, we
did manipulate whether the target's stealing behaviour was intentional or not 
(Game C), an approach which has been used in previous vignette studies to 
identify behaviour shaping motives [@Carlsmith2002]. The lower prevalence of 
behaviour shaping strategies in our study is consistent with prior work showing
that punishment often continues to be used when behaviour shaping is impossible,
such as when the target will never find out that they have been punished 
[@Crockett2014] or on the last round of repeated interactions [@Barclay2016]. We
found that demographic, personality, political, and religious variables tended
to be unrelated to behaviour shaping strategies, though this could reflect low
power to detect such associations given the low prevalence of these strategies
in our sample. We also found that participants accurately reported using the 
norm-enforcing strategy, but not the deterrent strategy. This finding is in line
with previous research showing that people struggle to accurately report the 
deterrent motivations for their punitive behaviour [@Carlsmith2002].

We defined antisocial punishment as occurring when a participant exclusively 
punished those who did not steal. There were no participants in our sample who 
followed this strategy, although some players did punish non-stealing
co-players. Harming non-stealing individuals was also consistent with the 
competitive strategies, which did appear in our sample albeit at low 
frequencies. The fact that no participants in our sample punished non-stealing 
across all games suggests that the antisocial punishment that has been observed
in other studies [@Herrmann2008; @Pleasant2018] is not aimed at harming 
cooperators specifically, as has been previously suggested [@Herrmann2008]. 
Instead, antisocial punishment is more likely to be motivated by improving one's 
relative position, which is in line with work showing that antisocial punishment 
disappears when relative payoffs cannot be changed (e.g., when punishment is 
only available with a 1:1 fee-fine ratio) [@Sylwester2013].

The fact that people use punishment for many different reasons poses problems 
for the way that punishment is operationalised in classic behavioural economic 
game studies. In these studies, a common assumption is that participants punish 
to change the behaviour of cheats [@Fehr2000; @Fehr2002]. But in reality, people 
may be choosing the punishment option to achieve a variety of different goals. 
This has implications for how people respond to being targeted by punishers in 
these games. Targets of punishment in these studies may know that punishment 
reflects different motives and can respond accordingly. For example, if targets
interpret punishment as serving a competitive motive, it may elicit retaliation
rather than encourage cooperation [@Dreber2008; @Nikiforakis2008b; 
@Raihani2019]. As punishers’ motives must be inferred (and such inferences
likely depend on character traits of the target, as well as the context in which 
punishment occurs), there is likely to be some variation and error in 
attributing motives to punishers. To the extent that inferred motives affect 
target responses, this might help to explain the mixed findings in the field as
to whether punishment actually motivates cheating targets to subsequently 
cooperate [@Raihani2019].

It is striking that the most common strategy in our dataset was to never punish. 
This is partly because punishment in these games imposes an economic cost for no 
tangible benefit. If the fee-fine ratio had been lower, such that it was cheaper 
to punish, we may have seen more punishment from participants. Indeed, 
`r papaja::printnum(mean(tar_read(dExc)$SelfRate10[tar_read(dExc)$strategy == "Never punish"] > 0, na.rm = TRUE)*100, digits = 0)`% 
of participants following the never punish strategy positively stated that they 
didn't want to pay to reduce anyone's bonus but would have done so if it were 
free. But the frequency of the never punish strategy perhaps also reflects a 
more general aversion to peer punishment, an aversion that has been highlighted 
in both WEIRD (Western, educated, industrialised, rich, and democratic) samples
[@Balafoutas2014; @Balafoutas2016] and in small-scale societies [@Baumard2010]. 
One reason that people may be averse to peer punishment is that, because it is 
a fundamentally harmful act, punishment can reflect badly on the punisher and 
people might therefore refrain from punishing others to avoid reputational 
damage [@Raihani2015]. People frequently avoid taking actions that could harm 
their reputation, even when they don’t know if reputation is at stake (as in the 
one-shot anonymous settings used here). Another reason that people may be averse
to peer punishment is that it can trigger retaliation [@Raihani2019]. This may 
be especially likely in situations that lack clear institutional norms to 
legitimise punishment, such as our economic games. As with reputation damage, 
people might abstain from peer punishment to avoid retaliation, regardless of 
whether retaliation is actually possible. By contrast, institutionalised 
punishment in small-scale societies often functions to compensate victims while 
limiting the potential for feuds and cycles of retaliation
[@Fitouchi2023; @Singh2022]. Future research should uncover whether people are 
more willing to punish in these conventionalised contexts (e.g., see ref
[@Molleman2019]).

One potential limitation with our design is that some strategies required more 
punishment than others, meaning that some strategies were more "expensive" to 
implement. For example, the competitive strategy required punishment in ten of 
twelve decisions, compared to the avoid disadvantageous inequity strategy which 
required only one instance of punishment (@tbl-strategies). We employed the 
strategy method to deal with this, calculating participant payoffs from a 
randomly chosen game instead of accumulating the costs across all games. This 
mitigates the concern that the total cost is the most salient feature 
determining participants' decisions. But even accounting for this, one could 
still argue that the difference in overall costs explains why, for example, the 
competitive strategy is less common in our dataset than the avoid 
disadvantageous inequity strategy. We do not think this feature of our design 
is a concern, however, for a number of reasons. First, we are interested in 
measuring strategies underlying *costly* punishment. Some of these strategies,
by their very nature, will be more costly to implement than others. This is 
reflected in our design. It would not make theoretical sense to use an 
alternative design where more or less punitive strategies are manipulated to 
cost participants the same amount. Second, when we asked participants whether
they would have punished if it were free to do so, we found low agreement 
with this statement (@suppfig-sliders1), suggesting that participants were not 
particularly sensitive to the costs of punishment. In line with this, when we 
plot the frequencies of different strategies against their expected costs, we 
find that cost is not a perfect predictor of strategy frequency: many "cheap" 
strategies are rare in our data and some "expensive" strategies are quite common
(e.g., always punish; @suppfig-spline). Third, this argument implies that the 
high frequency of the never punish strategy is merely an artefact of our design,
since it is the only strategy that does not cost anything to implement. But as
we have discussed, many other studies have found similar aversions to costly 
punishment in the lab [@Barclay2016; @Batistoni2022; @Bone2015; @Deutchman2021; 
@Raihani2012b] and in the real world [@Balafoutas2014; @Balafoutas2016; 
@Baumard2010] suggesting that this result is not an artefact of our design.

In sum, we have shown that while many people choose not to punish peers, those 
who do are motivated by a variety of different concerns, including behaviour 
shaping, egalitarianism, and competition. Much like the observed variation in 
human social learning strategies [@Molleman2014], humans thus also exhibit 
variation in their punishment strategies. These individual differences map onto 
personality dimensions, social preferences, political and religious views, and 
self-reports of behaviour. We hope that future work will continue to unpack the 
multifaceted nature of human punishment.

# Methods

## Ethical approval

This research was approved by the UCL Department of Psychology Ethics Committee 
(project: 3720/002) and ratified by the University of Auckland Human 
Participants Ethics Committee. The study was performed in accordance with all 
the relevant guidelines and regulations. Informed consent was obtained from all 
participants prior to the study and participation was voluntary.

## Pre-registration

We pre-registered the study on the Open Science Framework before collecting data 
in the United Kingdom (11^th^ November 2022; <https://osf.io/k75fc>). We 
submitted another pre-registration before collecting data in the United States 
(20^th^ June 2023; <https://osf.io/q4hdy>). In the pre-registrations, we 
outlined our study design, exclusion criteria, and analysis plan. As the study 
was exploratory, we did not pre-register any explicit hypotheses. We did not 
deviate from the pre-registrations.

## Exclusion criteria

We pre-registered that we would exclude participants who failed any of the 
attention checks, sped through the surveys (i.e., two standard deviations below 
the median duration), or flatlined (i.e., provided identical responses to matrix 
questions). We also stated that we would exclude data for particular games if 
participants failed the comprehension question for that game. We followed our 
pre-registered plan of conducting analyses with and without these exclusions 
(analyses without exclusions are reported in the Supplementary Material).

## Participants

```{r}
tar_load(c(d, dExc))
uk_n_full <- sum(d$Country == "United Kingdom")
uk_n_rtrn <- sum(d$Country == "United Kingdom" & !is.na(d$StartDate_Part2))
uk_n_excl <- sum(dExc$Country == "United Kingdom")
us_n_full <- sum(d$Country == "United States")
us_n_rtrn <- sum(d$Country == "United States" & !is.na(d$StartDate_Part2))
us_n_excl <- sum(dExc$Country == "United States")
```

We collected a representative sample of `r papaja::printnum(uk_n_full)` 
participants from the United Kingdom through the online platform Prolific 
(<https://www.prolific.com/>). All of these participants completed the economic 
games and `r papaja::printnum(uk_n_rtrn)` returned to complete the follow-up 
survey a week later 
(`r papaja::printnum((uk_n_rtrn / uk_n_full)*100, digits = 0)`% retention rate).
After exclusions, we were left with `r papaja::printnum(uk_n_excl)` participants 
overall (see @suppfig-sampleUK).

We later collected a representative sample of `r papaja::printnum(us_n_full)` 
participants from the United States through Prolific. All of these participants 
completed the economic games and `r papaja::printnum(us_n_rtrn)` returned to 
complete the follow-up survey 
(`r papaja::printnum((us_n_rtrn / us_n_full) * 100, digits = 0)`% retention 
rate). After exclusions, we were left with `r papaja::printnum(us_n_excl)` 
participants overall (see @suppfig-sampleUS for sample characteristics).

## Materials

### Economic games

In the first part of the study, participants completed six economic games, each 
with slight variations. In all games, there are multiple players and the 
participant takes the role of P1. P2 either (a) steals £0.20 from another player 
and adds it to their payoff or (b) does nothing. For each of these cases, 
participants are asked whether they would like to pay money to reduce P2's 
payoff. Games A-E have two players and Game F has three players.

The six games are as follows (variations bolded; see @fig-games for a visual 
representation of the games):

1. *Game A (Advantageous Inequity)*. P1 starts with £0.70 and P2 starts with 
   £0.10. P2 is given the option to either steal £0.20 from P1 or do nothing. P1 
   can then pay £0.10 to reduce P2’s payoff by £0.30.
2. *Game B (Equal)*. P1 starts with £0.70 and P2 starts with **£0.30**. P2 is 
   given the option to either steal £0.20 from P1 or do nothing. P1 can then pay 
   £0.10 to reduce P2’s payoff by £0.30.
3. *Game C (Computer)*. P1 starts with £0.70 and P2 starts with £0.30. 
   Participants are told that **"the computer will decide"** whether P2 steals
   £0.20 from P1 or does nothing. P1 can then pay £0.10 to reduce P2’s payoff by 
   £0.30.
4. *Game D (1:1 Fee-Fine)*. P1 starts with £0.70 and P2 starts with £0.30. P2 is 
   given the option to either steal £0.20 from P1 or do nothing. P1 can then pay 
   £0.10 to reduce P2’s payoff by **£0.10**.
5. *Game E (Disadvantageous Inequity)*. P1 starts with £0.70 and P2 starts with 
   **£0.50**. P2 is given the option to either steal £0.20 from P1 or do 
   nothing. P1 can then pay £0.10 to reduce P2’s payoff by £0.30.
6. *Game F (Third-Party)*. P1 starts with £1.00, P2 and P3 start with £0.70. P2 
   is given the option to either steal £0.20 **from P3** or do nothing. P1 can 
   then pay £0.10 to reduce P2’s payoff by £0.30.

<br>

We delineate nine punishment strategies that can be isolated based on players' 
decisions across the six economic games (@tbl-strategies). In existing 
literature, it is often unclear whether punishment decisions are sensitive to 
the previous actions of others or to payoff sensitivity. This is because (*i*) 
players' decisions typically introduce payoff differentials in economic games 
and (*ii*) where fine > fee, punishment tools can be used to harm co-players and 
to change relative payoffs [@Raihani2019]. Of the nine strategies we define, 
*deterrent*, *norm-enforcing*, *retributive* and *antisocial* are strategies 
that are only sensitive to the actions of others. Conversely, *avoid 
disadvantageous inequity*, *egalitarian*, *seek advantageous inequity* and 
*competitive* are strategies that are sensitive to payoffs. *Never punish* is an 
unconditional strategy (i.e., no sensitivity to actions or payoffs).

Our strategy definitions sometimes differ from how these strategies are defined 
and discussed in the literature (e.g., antisocial punishment). This is essential 
because previous definitions permit many possible motives, while we are aiming 
to isolate punishment motives. We define our strategies as follows:

- *Deterrent*: Deter others who have harmed you from harming you again in the 
future (when the target steals in Games A, B, D, and E). We note that while no 
behaviour change is possible in our one-shot games, deterrent motives can be 
isolated from retributive motives by asking whether participants are sensitive 
to the intentionality behind cheating, as in previous work [@Carlsmith2002]. 
- *Norm-enforcing*: Punish to enforce a shared anti-harm norm and encourage 
future norm compliance (when the target steals in Games A, B, D, E and F). 
Though we cannot measure future norm compliance in our one-shot games, these 
non-repeated (and 'stranger design') games have been used to identify 
norm-enforcing motives in many previous studies [@Fehr2000; @Fehr2002].
- *Retributive*: Punish if doing so harms another who harmed you (when the 
target steals in Games A-E). We note that retribution is also sometimes used to 
describe punishment of those who harmed others [@Carlsmith2002]. We require this 
tighter definition because our aim is to precisely identify punishment 
strategies, and sensitivity to whether self or others were harmed may be an 
important factor affecting the decision to punish (e.g., ref
[@Feldmanhall2014]). 
- *Avoid disadvantageous inequity*: Punish if doing so avoids disadvantageous 
inequity for the self (when the target steals in Game E). 
- *Egalitarian*: Punish if doing so makes payoffs for all more equal (when the 
target steals in Games E and F) (c.f. ref [@Dawes2007]).
- *Seek advantageous inequity*: Punish if doing so produces advantageous
inequity for the self (when the target steals in Games B and C). This is one 
possible motive underlying what others have previously called 'antisocial 
punishment' or 'spite'; here we isolate it.
- *Competitive*: Punish if doing so improves your relative position (all games 
except Game D). 
- *Antisocial*: Punish exclusively those who do not cause harm (punish 
non-stealing target in all games). This definition differs from some previous 
definitions of antisocial punishment, but this is essential to isolate the full 
range of motives that could be driving punishment and which are not typically 
disentangled in other settings. 
- *Never punish*: A person who never incurs a personal cost to punish a 
co-player (does not punish in any game).

<br>

For each game, participants saw the game instructions and answered a 
comprehension question before providing their decisions. After completing all 
the games, participants were asked to give an open-ended response explaining 
their behaviour in the games, and then responded to several slider questions 
capturing the different reasons for their decisions (for full wordings, see 
@supptbl-sliderWordings).

### Survey questions

In a follow-up survey, we collected the following data on participants (for 
wordings of all questions, see @supptbl-surveyWordings):

- *Demographics*. In the survey, we collected information on participants' 
  education level and self-reported socio-economic status (MacArthur ladder
  [@Adler2000]). We also collected additional demographic data from Prolific 
  (e.g., age, gender, student status).
- *Personality*. We used the Mini-IPIP scale [@Sibley2011] to measure the Big 6 
  personality dimensions of agreeableness ($\alpha$ = 
  `r papaja::printnum(tar_read(alphaAgree)$alpha)`), conscientiousness ($\alpha$ = 
  `r papaja::printnum(tar_read(alphaConsc)$alpha)`), extraversion ($\alpha$ = 
  `r papaja::printnum(tar_read(alphaExtra)$alpha)`), honesty-humility ($\alpha$ = 
  `r papaja::printnum(tar_read(alphaHonest)$alpha)`), openness to experience ($\alpha$ = 
  `r papaja::printnum(tar_read(alphaOpen)$alpha)`), and neuroticism ($\alpha$ = 
  `r papaja::printnum(tar_read(alphaNeur)$alpha)`). Four items were used for each 
  personality dimension.
- *Social Value Orientation*. We used the Social Value Orientation Slider 
  Measure to measure other-regarding preferences [@Murphy2011]. Across fifteen 
  items, participants made decisions on how to allocate different amounts of 
  money between themselves and another anonymous individual. From these 
  decisions, we calculated participants' Social Value Orientation "angle" as a 
  measure of their other-regarding preference, following the steps outlined in 
  ref [@Murphy2011].
- *Political ideology*. We included several measures of political ideology, 
  including left-right conservatism, Social Dominance Orientation [@Ho2015] 
  ($\alpha$ = `r papaja::printnum(tar_read(alphaSDO)$alpha)`; eight items), and 
  Right Wing Authoritarianism [@Bizumic2018] ($\alpha$ = 
  `r papaja::printnum(tar_read(alphaRWA)$alpha)`; six items). We also probed 
  participants' views on social inequality by asking them whether they would 
  like to bring people above (below) them on the MacArthur socio-economic status 
  ladder down (up) a peg or two.
- *Religious views*. We asked participants how religious they consider 
  themselves and whether they believe that God or another spiritual non-human 
  entity controls the events in the world [@Laurin2012].

<br>

## Procedure

We began data collection in the United Kingdom on 28^th^ November 2022, with 
participants returning to complete the follow-up survey on 5^th^ December 2022. 
We then ran a second wave of data collection in the United States on 20^th^ June 
2023, with participants returning to complete the follow-up survey on 27^th^ 
June 2023. Our surveys were designed through the online survey platform 
Qualtrics (<https://www.qualtrics.com/>).

In the initial games survey, participants completed all six economic games in a 
random order, with punishment decisions (whether to punish a stealing target and 
whether to punish a target who did nothing) randomised within games. Responses 
to comprehension questions suggested that participants understood the six 
economic games (@supptbl-comp). In order to partially mitigate the fact that 
some punishment strategies predict more punishment than others and are thus more 
"expensive" to implement (@tbl-strategies), we used the strategy method to 
incentivise the economic games, choosing a random game to determine bonus
payments rather than summing participants' earnings across all games. After all
games, 
`r papaja::printnum((sum(dExc$DecisionConsequences == "Yes", na.rm = TRUE) / nrow(dExc)) * 100, digits = 0)`% 
of participants stated that they believed that their decisions had real 
consequences for others.

In the follow-up survey, participants completed blocks of questions on 
demographics, personality, Social Value Orientation, political ideology, and 
religious views in a random order, with questions randomised within blocks. A 
random decision from the Social Value Orientation Slider Measure was chosen to 
determine bonus payment.

Participants were paid £1.80 for completing the games survey, plus a bonus 
payment from the six economic games (between £0.40 -- £0.70 depending on their 
decisions). Participants were paid £1.50 for completing the follow-up survey, 
plus a bonus payment from the Social Value Orientation Slider Measure (between 
£0.50 -- £0.85 depending on their decisions).

## Statistical analysis

We pre-registered that we would use a Bayesian latent state model to infer 
unobserved punishment strategies from the observed data (for a similar version 
of this model, see ref [@McElreath2020]). In this model, participants $i$ in 
countries $c$ make binary punishment decisions across twelve decisions $j$. We 
assume that the probability of the observed data $y_{i,j}$ is the weighted 
average of the probability of the observed data conditional on each of the ten 
punishment strategies $s$. From this logic, the model estimates the probability 
of each strategy $p_{s}$. The full model is as follows:

$$
\begin{align}
y_{i,j} &\sim \text{Bernoulli}(\theta_{j}) \\
\theta_{j} &= \sum_{s=1}^{10} p_{s} \text{Pr}(\text{punish}|s,j) \nonumber \\
p &= \text{softmax}(\alpha_{c[i]}) \nonumber \\
\alpha_{s,c} &\sim \text{Normal(0, 1)} \nonumber
\end{align}
$$

```{r}
error <- tar_read(post5)$error
```

The conditional probabilities $\text{Pr}(\text{punish}|s,j)$ are hard coded in 
the model as outlined in @tbl-strategies. We incorporate an implementation error
rate $\delta$ into these conditional probabilities by coding ticks in 
@tbl-strategies with a conditional probability of $1 - \delta$ and coding 
crosses with a conditional probability of $0 + \delta$. We set $\delta$ to 0.05
in all models, which is similar to its value when we estimate it as a free 
parameter in an additional model (median posterior $\delta$ = 
`r papaja::printnum(median(error))`, 95% CI
[`r papaja::printnum(quantile(error, 0.025))`
`r papaja::printnum(quantile(error, 0.975))`]; @suppfig-model5). The random 
choice strategy is consistently coded with a conditional probability of ½ across
all decisions.

To include a categorical predictor in the model, we estimate a different 
$\alpha_{s,c}$ for each categorical level. To include a continuous predictor 
$x$ in the model, we include a slope $\beta$ in the linear model for $p$:

$$
\begin{align}
y_{i,j} &\sim \text{Bernoulli}(\theta_{j}) \\
\theta_{j} &= \sum_{s=1}^{10} p_{s} \text{Pr}(\text{punish}|s,j) \nonumber \\
p &= \text{softmax}(\alpha_{c[i]} + \beta_{c[i]}x_{i}) \nonumber \\
\alpha_{s,c} &\sim \text{Normal(0, 1)} \nonumber \\
\beta_{s,c} &\sim \text{Normal(0, 0.2)} \nonumber
\end{align}
$$

These models control for multiple comparisons across strategies by estimating 
the effects of the predictor on all strategies simultaneously.

We estimated the posterior distributions of these models using Hamiltonian Monte 
Carlo as implemented in Stan version 2.32.2 [@Stan2020]. We ran each model for 
2000 samples, with 1000 warmup samples. R-hat values and effective sample sizes 
suggested that all models converged normally. Trace plots are reported in 
@suppfig-trace.

We validated the model by simulating observed data (_n_ = 100) from a known 
frequency of strategies. The model was successfully able to recover the known 
frequency of strategies from the simulated data (@suppfig-sim).

## Reproducibility

All data and code are accessible on GitHub:
<https://github.com/ScottClaessens/punishStrategies>. All analyses were 
conducted in R version 4.4.2 [@RCoreTeam]. Visualisations were created with the
*ggplot2* [@Wickham2016] and *cowplot* [@Wilke2020] R packages. We used the
*targets* [@Landau2021] R package and *quarto* [@Allaire2024] to reproducibly 
generate the manuscript.

\newpage

# Acknowledgements

This work was supported by a Royal Society of New Zealand Catalyst Leaders Grant 
to Q.D.A and N.R. (ref: ILFUOA2002).

# Competing Interests

The authors declare no competing interests.

# Data Availability

All data used in this study are publicly available on GitHub: 
<https://github.com/ScottClaessens/punishStrategies>

# Code Availability

All code to reproduce the analyses in this study are publicly available on 
GitHub: <https://github.com/ScottClaessens/punishStrategies>

\newpage

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

\newpage

{{< include appendix.qmd >}}